### 2016-01-22 manuscript/f1000_submission/All.tex
closer to submission ready for f1000

%latex.default(All(mc_chla_tbl))%
\begin{table}[!tbp]
\begin{center}
\begin{tabular}{l}
\hline\hline
\multicolumn{1}{c}{}\tabularnewline
\hline
+\tabularnewline
Heading(`Cond. Probability`) * c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9) + Heading(`USEPA Child`) * c(0.07, 0.07, 2.98, 10.67, 23.36, 38.3, 65.6, 117.5, 167.04) + Heading(`WHO Drink`) * c(0.07, 4.43, 20.43, 42.24, 66.96, 103.2, 166.63, 338.4, 516) + Heading(`USEPA Adult`) * c(0.07, 11.8, 31.4, 66.96, 83.52, 133.2, 871.2, 871.2, 871.2)\tabularnewline
Heading(`WHO Recreational`) * c(1.22, 20, 53.64, 82.22, 105.84, 155.52, 216, 871.2, 871.2)\tabularnewline
\hline
\end{tabular}\end{center}

\end{table}